---
title: "Store Sales Forecasting"
subtitle: "Kaggle Competition"
format: 
  html:
    toc: true
    embed-resources: true
author: "Ivan Morales, Rafael Takata"
---

# Initial setup

## Packages

```{r}
#| label: pkgs
#| message: false
#| warning: false

library(conflicted)
library(tidyverse)
library(tsibble)
library(feasts)
library(fable)

conflicts_prefer(
  lubridate::date(),
  dplyr::filter()
)
```

# Data import

## `train.csv`

### Convert to a `tsibble`

```{r}
#| label: train
train_tsbl <- read_csv("data/train.csv", 
                       show_col_types = FALSE) |> 
  as_tsibble(index = date, key = c(store_nbr, family)) |> 
  select(-onpromotion)

train_tsbl
```

### Checking for gaps

```{r}
train_tsbl |> 
  has_gaps()
```

All of them have implicit gaps in time. Below we'll find where are such gaps:

```{r}
gaps <- train_tsbl |> 
  count_gaps(.full = TRUE) |> 
  distinct(.from)

gaps
```

It's Christmas. We'll fill the implicit gaps and set to zero its value.

```{r}
#| label: fill_gaps
train_tsbl <- train_tsbl |> 
  fill_gaps(.full = TRUE, 
            sales = 0L)

train_tsbl
```

## additional data

### `holidays_events.csv`

```{r}
holidays <- read_csv("data/holidays_events.csv", show_col_types = FALSE)
holidays
```
### `transactions.csv``

```{r}
transactions <- read_csv("data/transactions.csv", show_col_types = FALSE)
transactions
```

### `stores.csv`

```{r}
stores <- read_csv("data/stores.csv", show_col_types = FALSE)
stores
```

### `oil.csv`

```{r}
oil <- read_csv("data/oil.csv", show_col_types = FALSE) |> 
  as_tsibble(index = date) |> 
  fill_gaps(.full = TRUE) |> 
  fill(dcoilwtico, .direction = "downup")
oil 
```

### Joining the data in one `tsibble`

```{r}
holidays <- holidays |> 
  filter(transferred == FALSE) |> 
  mutate(
    feriado_local = ifelse(locale == "Local", 1, 0),
    feriado_regional = ifelse(locale == "Regional", 1, 0),
    feriado_nacional = ifelse(locale == "National", 1, 0)
  )
holidays
```
  
```{r}

```


```{r}
train_tsbl <- train_tsbl |> 
  left_join(oil, by = "date") |> 
  left_join(stores, by = "store_nbr")
train_tsbl
```

### `test.csv`

```{r}
test <- read_csv("data/test.csv", show_col_types = FALSE) |> 
  select(-onpromotion)
test
```

# Exploratory analysis

## Products with no sales

We'll check if there are any family products not being sold in specific stores:

```{r}
ceros <- train_tsbl |> 
  as_tibble() |> 
  group_by(store_nbr, family) |> 
  summarise(sales = sum(sales), .groups = "drop") |> 
  filter(sales == 0) |> 
  select(-sales)

ceros
```

There are `{r} nrow(ceros)` series with no sales in the training set. We'll use a **NAIVE** model to forecast them.The time series that do have sales will be stored in `series_tsbl`:

```{r}
series_tsbl <- train_tsbl |> 
  anti_join(ceros, by = c("store_nbr", "family"))

series_tsbl
```

and the ones without sales in `series_ceros`:

```{r}
series_ceros <- train_tsbl |> 
  anti_join(series_tsbl, by = c("date", "store_nbr", "family"))
series_ceros
```

The forecasts for such series are done below:

```{r}
series_ceros_fit <- series_ceros |> 
  model(naive = NAIVE(sales))
series_ceros_fit

series_ceros_fcst <- series_ceros_fit |> 
  forecast(h = 16)

series_ceros_fcst
```

# Benchmark model

## Fit

```{r}
tictoc::tic()
series_tsbl_fit <- series_tsbl |> 
  model(
    ETS = ETS(sales)
  )
tictoc::toc()

series_tsbl_fit
```

```{r}
tictoc::tic()
accuracy(series_tsbl_fit)
tictoc::toc()
```

## Forecast

```{r}
tictoc::tic()
series_tsbl_fcst <- series_tsbl_fit |> 
  forecast(h = 16)
tictoc::toc()

series_tsbl_fcst
```

Joining forecasts

```{r}
full_fcst <- series_tsbl_fcst |> 
  bind_rows(series_ceros_fcst)

full_fcst
```

## Submission

```{r}
submission <- full_fcst |> 
  left_join(test, by = c("date", "store_nbr", "family")) |> 
  as_tibble() |>
  select(id, .mean) |>
  rename(sales = .mean)

submission
```

```{r}
#submission |> 
  #write_csv("submission1_p25.csv")
```
